Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs

### 2015-07-06 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt

We would like to thank the reviewers for their detailed and constructive feedback. We believe we have addressed
all the reviewer comments (or explained below why we did not implement a suggestion) and believe this
has made a stronger paper. Significant text changes are highlighted in yellow in the paper, and 
we summarize the changes made to address each comment or answer any detailed questions in the response
below.

Reviewer: 1

The paper presents significant space talking about how to transform a hyper graph to a graph using various models.  This has been well studied in the past.  It would be useful to present the proposed techniques in the context of this previous work.  As a start, there are numerous references within reference [22] that should be considered.  Doing so would help the reader understand the novelty of the hypergraph->graph methodology and results.

[Javeed to add survey paper reference and discussion on why we use a non-dummy node star topology][DONE]

@@

In section IV.A, you should mention that VPR's placer is simulated annealing-based. In particular, the reason you cited in the last paragraph of IV.A is based on the assumption that you are using an iterative improvement-based placer (and not, say, an analytical placer).

[Ehsan: add mention of annealing][DONE]

For the discussion about hypergraph to graph transformation in Sec V.C, there were many papers discussing how to set the edge weights in the 90's, please cite them (for example, see the survey paper "Recent directions in netlist partioning: a survey"). In particular, setting the edge weight to 1/(n-1) instead of 1/n is a popular choice since it can model a 2-pin net exactly. So, it makes sense to report the results for 1/(n-1). It looks like your star model picks one node of a net and connect it to other nodes of the net, but the more popular star model in the physical design literature would add a dummy node for each net that connects to all nodes of the net. So please explicitly define your star model in words. And from p.9, it looks like that you also want to distinguish the source of a net from the sinks of a net in the star model, if so, please state that in the definition of your star model.

[Javeed: make explicit and correct text if it says we make all hyperedges the same weight with 1/n (they will vary from 1/2 to 1 in total hyperedge weight).] [DONE]

@@

(1) The case for METIS over hMETIS is weak

Both METIS and hMETIS are well known tools for graph and hypergraph partitioning, respectively. Since circuits are known to be hypergraphs, the natural inclination would be to use hMETIS; however, the authors argue that METIS is more appropriate because hMETIS does not support heterogeneous balance constraints. Therefore, the authors take great pains to convert the hypergraph to a graph.

This is not satisfying to me at all. If the correct solution is to use hypergraph partitioning with heterogeneous balance constraints, then the authors should make appropriate modifications to hMETIS to produce the appropriate partitioning tool. It is unclear how much relevant information is lost by transforming the hypergraph to a graph, and what the impact on relative CAD metrics (W_min, f_max, etc) would be.

[Javeed: mention hMetis is not open source (emailed person), give results on hmetis vs. metis cut size (with and without balance), and mention the star model should help timing, as timing is a connection / edge, phenomenon, not a net / hyperedge phenomenon.] [DONE]
The hMetis source code is not freely available. We contacted the authors of hMetis, who stated that an hMetis update is in development and will contain support for heterogeneous balance constraints. Our experiments showed that hMetis performed 30% better on average when compared to Metis on our benchmark circuits, when both were used with homogeneous balance constraints.

### 2015-07-06 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Minor revisions in response to reviewer comments.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4674 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

The current version of hMetis does not support heterogeneous balance constraints, though Metis does, so we use Metis in our CAD flow. However, as the circuit netlist is naturally represented as a hypergraph, it needs to be transformed to a graph before Metis can process it.

@@

We explored several ways to generate the per-net subgraphs, varying the graph topology and the edge weight scheme. We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}. \hl{Ref.\mbox{\cite{hypergraph_to_graph_survey}}
considers the addition of ``dummy'' nodes to model a hypergraph as a graph, with the objective of obtaining equivalent edge cut weights after partitioning. In our star model, we preferred to designate the net source as the center of the star, rather than introduce a dummy node. This choice was motivated by a desire to optimize timing -- the star topology clearly differentiates sources from sinks, and we expect this to guide the partitioner to keep sources and sinks in the same partition.}

@@

The performance of the clique topology is significantly worse than that of the star topology, regardless of edge weight scheme. An intuitive explanation for this result is that with the clique topology, the partitioner does not have knowledge of which vertex is the source of a net and which vertices are sinks. In contrast, the star topology clearly differentiates sources from sinks and this appears to give the partitioner an “anchor point” that pulls all hyperedge fanouts toward the source. Additionally, if $n$ is the average number of vertices in a hyperedge, the clique topology generates $O(n^2)$ edges while the star topology generates only $O(n)$ edges, so the star topology saves both memory and runtime.

For both topologies, the $1/n$ edge weight scheme gives the smallest cut size on average. This can be explained intuitively in terms of the total edge weight over all edges of the graph. The star $1$ (constant) edge weight scheme assigns a total weight of $n$ to each net, which heavily weights high-fanout nets. In contrast, the star $1/n$ model assigns the same total weight of $1$ to each net. The star $1/n^2$ model penalizes high-fanout nets relative to lower-fanout nets. Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ model achieves this.

@@

The star topology with 1/n edge weights achieves the best minimum channel width, confirming that hyperedge cutsize is a good proxy for routability. Consequently, we use this model in all future results in this paper. Unlike routability, we found that the post-routing critical path delay was not strongly impacted by the hypergraph model used.

>>>>

The current version of hMetis does not support heterogeneous balance constraints, though Metis does, so we use Metis in our CAD flow. However, as the circuit netlist is naturally represented as a hypergraph, it needs to be transformed to a graph before Metis can process it. \hl{Our experiments showed that hMetis performed 30\% better on average when compared to Metis on our benchmark circuits with homogeneous balance constraints.}

@@

We explored several ways to generate the per-net subgraphs, varying the graph topology and the edge weight scheme. We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}. \hl{Ref.\mbox{\cite{hypergraph_to_graph_survey}} surveys several methods of transforming hypergraphs to graphs, and considers the addition of ``dummy'' nodes to the star model. In our star model, we preferred to designate the net source as the center of the star, rather than introduce a dummy node. This choice was motivated by a desire to optimize timing -- the star topology clearly differentiates sources from sinks, and we expect this to guide the partitioner to keep sources and sinks in the same partition.}

@@

The performance of the clique topology is significantly worse than that of the star topology, regardless of edge weight scheme. An intuitive explanation for this result is that with the clique topology, the partitioner does not have knowledge of which vertex is the source of a net and which vertices are sinks. In contrast, the star topology clearly differentiates sources from sinks and this appears to give the \hl{iterative refinement based} partitioner an “anchor point” that pulls all hyperedge fanouts toward the source. Additionally, if $n$ is the average number of vertices in a hyperedge, the clique topology generates $O(n^2)$ edges while the star topology generates only $O(n)$ edges, so the star topology saves both memory and runtime.

\hl{For both topologies, the $1/n$ edge weight scheme gives the smallest cut size on average. This can be explained intuitively in terms of the total edge weight over all edges of the graph. The star $1$ (constant) edge weight scheme assigns a total weight of $n - 1$ to each net, which heavily weights high-fanout nets. In contrast, the star $1/n$ model assigns a total weight of $1 - \frac{1}{n}$ to each net. The star $1/n^2$ model penalizes high-fanout nets relative to lower-fanout nets. Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ model best achieves this.}

@@

The star topology with $1/n$ edge weights achieves the best minimum channel width, confirming that hyperedge cutsize is a good proxy for routability. Consequently, we use this model in all future results in this paper. Unlike routability, we found that the post-routing critical path delay was not strongly impacted by the hypergraph model used.
